%!TEX root=../main.tex
\begin{abstract}
%put this line in intro. Abstract is getting too big.
Despite having an advertised arrival time,
a flights' actual arrival time can be far from the scheduled time.
One metric, arrival status, can be used to measure the disparity between
the scheduled arrival time and the actual arrival time of a flight.
Arrival status of a flight can be early, on-time or delayed.
In this project, our goal was to predict the arrival status of a flight 1-3 days earlier than the actual flight.
%To reduce the scale of the experiment, 
Solving the general problem of predicting arrival status for any flight is too large to solve within the time-frame of one month.  
Hence, we attempted to solve a miniature version of the problem where we build a model to predict the arrival status of 6 flights arriving at Syracuse from 3 different origins (New York, Chicago, and Orlando).
Given the miniature size of our problem, we did not employ any deep learning techniques to built the model. 
We built a tree-based based machine learning model from the historical data of flights
that were operating between 2006 and 2024.
In our model, we incorporated 3 categories of flight data: flight characteristics, airlines on-time arrival history, and airport's on-time departure history. 
Additionally, we incorporated historical weather data in our model.
We built an ensemble of trees by using XGBoost algorithm to predict the arrival status of a flight given the flight date, flight no, origin, scheduled departure time, and scheduled arrival time.
We evaluated our model based on its' prediction accuracy on the test dataset and
based on its' accuracy in predicting arrival status of future flights operated in the period of 4 days (from April 19, 2024 to April 23, 2024).
On our test dataset of 2818 samples, our model achieves 52.7\% accuracy, while predicting arrival status of future flights, our model were correct on 12 out of 23 occasions.
%Due to time constraint, we tried to optimize our model for predicting arrival status of
%6 flights arriving at Syracuse from 3 different origins (New York, Chicago and Orlando).  
%John F. Kennedy International Airport (JFK), Chicago O'Hare International Airport (ORD), and 
\\
\end{abstract}
