%!TEX root=../main.tex

\section{Problem Description}
\label{sec:problem}

Aviation industry tries to follow pre-calculated schedule for each step of their operation.
Airlines mention schedule departure time and scheduled arrival time of a flight while they are selling tickets for that flight.
But while operating the flight, actual departure or arrival time might be different from the time scheduled earlier. 
In this project, we are more concerned about the arrival time of a flight.

Depending on the amount of difference between the scheduled arrival time and actual arrival time,
we are classifying arrival status of a flight as ``EARLY'', ``ON-TIME'', or ``LATE.''
If actual arrival time of a flight is falls between the 5 minutes offset (5 minutes earlier or 5 minutes later) of the scheduled arrival time,
we are classifying the arrival status as ``ON-TIME.'' 
If a flight arrives more than 5 minutes earlier than the scheduled time, we classify the arrival status of the flight as ``EARLY.''
On the other hand, if it arrives more than 5 minutes later than the scheduled time, we classify the arrival status of the flight as ``LATE.''

Now, solving the general problem (predicting arrival status of all operating flights) would require way too much data
and development of sophisticated machine learning algorithms.
Given the one-month time period and the restriction on using neural network based algorithms, we decided to solve a relatively specific version of the problem.

In this version, instead of predicting arrival status of all flights, we decided to predict arrival status of 6 flights at one specific airport, Syracuse Hancock International Airport (SYR), from 3 different origin airports:
John F. Kennedy International Airport (JFK),
O'Hare International Airport (ORD), and Orlando International Airport (MCO).
In particular, Six flights (3 pairs from each origin) that we choose for predicting arrival status are: United Airlines UA 1400 and American Airlines AA 3402 from ORD in Chicago,
Jet Blue B6 116 and Delta Airlines DL 5182 from JFK in New York, 
and Jet Blue B6 656 and Southwest Airlines WN 5285 from MCO in Orlando. 
Flights in each pair from the same origin departs in sequence.
For example, from ORD, UA 1400 departs at 18:52 and AA 3402 departs at 19:59.
First, we need to predict arrival status (at SYR) of the earlier flight (UA 1400).
After that, given the arrival status (EARLY, ON-TIME, LATE) of the earlier flight, we need to predict the arrival status of the later flight (AA 3402).
For the first prediction, the prediction query will contain flight date, flight no, origin, scheduled departure time, and  scheduled arrival time. 
For the second prediction, prediction will query will contain an additional information which is the arrival status of the earlier flight.
 